---
title: "Diabetes Prevalence by State"
description: A comparison on the prevalence of diabetes among adults by state to rates of depression amongst adults with diabetes and mortality rates
site: distill::distill_website
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

# Learn more about creating websites with Distill at:
# https://rstudio.github.io/distill/website.html

# Learn more about publishing to GitHub Pages at:
# https://rstudio.github.io/distill/publish_website.html#github-pages
library(tidyverse)
library(readxl)
library(rmarkdown)


```

```{r, include = FALSE, echo = FALSE}
CDI_data <- read_csv("CDI.csv")

```
```{r, echo = FALSE, layout = "1-page"}
clean_d_data <- CDI_data %>% 
  filter(Question == "Prevalence of diagnosed diabetes among adults aged >= 18 years", Stratification1 == "Overall", DataValueType == "Age-adjusted Prevalence", YearStart %in% c(2013:2018)) %>% 
  drop_na(DataValueAlt) %>% 
  mutate(region = tolower(LocationDesc)) %>% 
  select(YearStart, region, Question, DataValueAlt, DataValueUnit)

map_data(map = "state") %>% 
  left_join(clean_d_data, by = "region") %>% 
  ggplot(mapping = aes(x = long, y = lat, group = group, fill = DataValueAlt)) +
  geom_polygon() +
  facet_wrap(~YearStart) +
  labs(title = "Prevalence of Diagnosed Diabetes among Adults by State",
       subtitle = "Higher prevalence of diabetes is seen in the southern region of the US",
       fill = "Age-adjusted\nPrevalence (%)",
       x = "Longitude",
       y = "Latitude") +
  theme_classic()
```

As can be seen from the above plot, the southern region of the United States appears to continuously have higher rates of diabetes than other regions. To contrast, the northwestern region has continously lower rates of diabetes. However, these trends seen above do not continue to the plot below.

```{r, layout = "1-page"}
clean_dd_data <- CDI_data %>% 
  filter(Question == "Prevalence of depressive disorders among adults aged >= 18 years with diagnosed diabetes", 
         Stratification1 == "Overall", 
         DataValueType == "Age-adjusted Prevalence",
         YearStart %in% c(2013:2018)) %>% 
  drop_na(DataValueAlt) %>% 
  mutate(region = tolower(LocationDesc)) %>% 
  select(YearStart, region, Question, DataValueAlt, DataValueUnit)


map_data(map = "state") %>% 
  left_join(clean_dd_data, by = "region") %>% 
  ggplot(mapping = aes(x = long, y = lat, group = group, fill = DataValueAlt)) +
  geom_polygon() +
  facet_wrap(~YearStart) +
  labs(title = "Prevalence of Depressive Disorders among Adults Diagnosed\nwith Diabetes by State",
       subtitle = "Lower prevalence of depressive disorders among adults with diabetes\nis most consistently seen in California",
       fill = "Age-adjusted\nPrevalence (%)",
       x = "Longitude",
       y = "Latitude") + 
  theme_classic() 
```
While rates of diabetes appear to be highest in the southeastern region of the United States, rates of depressive disorders among adults with diabetes has more variation, with the most constant trend being lower rates in California and in some states in the northeastern region.

```{r, layout = "1-page"}
clean_diabetes_data <- CDI_data %>% 
  mutate(diabetes = ifelse(str_detect(Question, "diabetes"), TRUE, FALSE)) %>% 
  filter(diabetes == TRUE,
         DataValueType == "Age-adjusted Rate",
         Stratification1 == "Overall",
         Question == "Mortality due to diabetes reported as any listed cause of death",
         YearStart %in% c(2013:2018)) %>% 
  drop_na(DataValueAlt) %>% 
  mutate(region = tolower(LocationDesc)) %>% 
  select(YearStart, region, Question, DataValueAlt, DataValueUnit) %>% 
  mutate(DataValueAlt = DataValueAlt/1000)

map_data(map = "state") %>% 
  left_join(clean_diabetes_data, by = "region") %>% 
  ggplot(mapping = aes(x = long, y = lat, group = group, fill = DataValueAlt)) +
  geom_polygon() +
  facet_wrap(~YearStart) +
  labs(title = "Mortality Rate due to Diabetes",
       subtitle = "Oklahoma consistently has the highest rates of mortality due to diabetes",
       fill = "Age-adjusted\nRate (%)",
       x = "Longitude",
       y = "Latitude") + 
  theme_classic()
  
```
As can be seen in the above plot, most states have a lower mortality rate due to diabetes, which also doesn't follow the trends seen in either of the above plots. However, it is interesting to note the higher rate in Oklahoma as compared to all other states since there isn't a significantly higher rate of diabetes in this state. 
